How to select predictive models for decision-making or causal inference?

### 2024-01-10 how_to_select_causal_models.tex
iter
add graphical abstract
first proposition of reviewers
fix jamia demand
fix for submission

%\usepackage[a4paper, left=15mm, right=15mm, top=20mm,]{geometry}
% XXX: for submission
\usepackage[a4paper, left=25mm, right=25mm, top=20mm,]{geometry}
% declared functions and additional packages
\usepackage{caption}
\usepackage{subcaption}
\usepackage{threeparttable}
\usepackage{times}
\usepackage{hyperref}
\def\mycitecolor{green!50!black}
\def\mylinkcolor{red!60!black}
\hypersetup{ colorlinks=true, linkcolor=\mylinkcolor, citecolor=\mycitecolor }
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb} %mathbb packages
\usepackage{siunitx}
\usepackage{algorithm, algorithmicx, algpseudocode}
\usepackage{multirow}
\usepackage[sorting=nyt,style=numeric,natbib=true,doi=false,isbn=false,url=false,eprint=false, maxnames=999, maxcitenames=2]{biblatex} 
% To cite with superscript
\let\cite=\supercite
\DeclareNameAlias{default}{last-first}
\usepackage{placeins}
% associated arxiv class
%\usepackage{arxiv}
\usepackage{multicol}
\setlength{\columnsep}{0.5cm}
\usepackage{wrapfig}
%\usepackage{wraptable}
\usepackage{authblk}% for author affiliations
\usepackage[dvipsnames]{xcolor}

@@

\author[1,2,*]{Matthieu Doutreligne}
\author[1]{Gaël Varoquaux}

@@

% XXX: for submission
%\twocolumn
% XXX: for submission
\setstretch{1.1}
\section{Introduction}\label{sec:intro}

@@

We first illustrate the problem of causal model selection and briefly review
prior art. Then we anchor causal model selection in the \emph{potential outcome}
framework and details the causal risks and model-selection procedure. We then
rewrite the so-called $R\text{-risk}$ as a reweighted version of mean squared
difference between the true and estimated individualized treatment effect.
Finally, we conduct a thorough empirical study comparing the different metrics
on diverse datasets, using a family of simulations and real health
data, going beyond prior work limited to specific simulation settings
\cite{schuler_comparison_2018,alaa_validating_2019} (Appendix
\ref{apd:prior_work}).

@@

\section{Discussion}\label{sec:discussion}

@@

\section{Conclusion}\label{sec:conclusion}
\paragraph{More $R\text{-risk}$ to select models driving decisions}

@@

\appendix
%\setcounter{secnumdepth}{0}
\setcounter{secnumdepth}{3}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{images/2023-03-08-11-10-28_acic_2016_ate_heterogeneity.parquet_abs_bias_ylog_scale=True.pdf}%
    \caption{\textbf{Different outcome models lead to different
            estimation errors on the Average Treatment Effects},
        on 77 classic simulations with known true causal effect
        \cite{dorie_automated_2019}. The different models are ridge regression
        and random forests with different hyper-parameters
        (details
        \ref{apd:toy_example:acic_2016_ate_variability}). The different configurations are
        plotted as a function of increasing difference between treated and
        untreated population --see
        \autoref{subsec:measuring_overlap}.
        There is no systematic best performer; data-driven model
        selection is important.
        \label{fig:acic_2016_ate_heterogeneity}%
    }
\end{figure}

@@

% \begin{figure}[htbp]
%   \centering
%   \caption{Caussim simulations (500 repetitions):
%   $R\text{-risk}_{IPW^*_{naive}}$ (in green) is exploding}\label
%   {apd:simu_noised}
%   \includegraphics[width=0.7\linewidth]{images/noised_0.001_simu_best_normalized_abs_bias_ate_vs_mmd.pdf}
% \end{figure}

>>>>

%\usepackage[a4paper, left=15mm, right=15mm, top=20mm,]{geometry}
% XXX: for submission
\usepackage[a4paper, left=25mm, right=25mm, top=20mm,]{geometry}
% declared functions and additional packages
\usepackage{caption}
\usepackage{subcaption}
\usepackage{threeparttable}
\usepackage{times}
\usepackage{hyperref}
\def\mycitecolor{green!50!black}
\def\mylinkcolor{red!60!black}
\hypersetup{ colorlinks=true, linkcolor=\mylinkcolor, citecolor=\mycitecolor }
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb} %mathbb packages
\usepackage{siunitx}
\usepackage{algorithm, algorithmicx, algpseudocode}
\usepackage{multirow}
\usepackage{appendix}
\usepackage[sorting=nyt,style=numeric,natbib=true,doi=false,isbn=false,url=false,eprint=false, maxnames=999, maxcitenames=2]{biblatex} 
% To cite with superscript
\let\cite=\supercite
\DeclareNameAlias{default}{last-first}
\usepackage{placeins}
% associated arxiv class
%\usepackage{arxiv}
\usepackage{multicol}
\setlength{\columnsep}{0.5cm}
\usepackage{wrapfig}
%\usepackage{wraptable}
\usepackage{authblk}% for author affiliations
\usepackage[dvipsnames]{xcolor}
\usepackage{setspace} % double spaced required by journal
\doublespacing

@@

\author[1,2,*]{Matthieu Doutreligne, PhD}
\author[1]{Gaël Varoquaux, PhD}

@@

% XXX: for submission
%\twocolumn
% XXX: for submission (double spaced required)
%\setstretch{1.1}
\doublespacing
\section{Introduction}\label{sec:intro}

@@

We first illustrate the problem of causal model selection. Then we anchor causal model selection in the \emph{potential outcome}
framework and details the causal risks and model-selection procedure. We then
rewrite the so-called $R\text{-risk}$ as a reweighted version of mean squared
difference between the true and estimated individualized treatment effect.
Finally, we conduct a thorough empirical study comparing the different metrics
on diverse datasets, using a family of simulations and real health
data, going beyond prior work limited to specific simulation settings
\cite{schuler_comparison_2018,alaa_validating_2019} (Appendix
\ref{apd:prior_work}).

@@

\section{Discussion and conclusion}\label{sec:discussion}\label{sec:conclusion}

@@

\paragraph{More $R\text{-risk}$ to select models driving decisions}

@@

\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\thetable{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}

\begin{appendices}

    \part*{Supplementary materials:\\
        How to select predictive models for decision-making or causal inference? }

    \setcounter{secnumdepth}{3}

    %\part{Supplementary materials}

@@

    \begin{figure}[!b]
        \centering
        \includegraphics[width=0.95\linewidth]{images/2023-03-08-11-10-28_acic_2016_ate_heterogeneity.parquet_abs_bias_ylog_scale=True.pdf}%
        \caption{\textbf{Different outcome models lead to different
                estimation errors on the Average Treatment Effects},
            on 77 classic simulations with known true causal effect
            \cite{dorie_automated_2019}. The different models are ridge regression
            and random forests with different hyper-parameters
            (details
            \ref{apd:toy_example:acic_2016_ate_variability}). The different configurations are
            plotted as a function of increasing difference between treated and
            untreated population --see
            \autoref{subsec:measuring_overlap}.
            There is no systematic best performer; data-driven model
            selection is important.
            \label{fig:acic_2016_ate_heterogeneity}%
        }
    \end{figure}

@@

    % \begin{figure}[htbp]
    %   \centering
    %   \caption{Caussim simulations (500 repetitions):
    %   $R\text{-risk}_{IPW^*_{naive}}$ (in green) is exploding}\label
    %   {apd:simu_noised}
    %   \includegraphics[width=0.7\linewidth]{images/noised_0.001_simu_best_normalized_abs_bias_ate_vs_mmd.pdf}
    % \end{figure}
\end{appendices}
